---
title: "visualization"
author: "Yijia"
date: "2026-10-01"
output: github_document
---

```{r}
library(tidyverse)
library(ggridges)
library(patchwork)
library(p8105.datasets)
data("weather_df")
```

Now we have everything we need
```{r}
weather_df
```

Let's make a scatterplot
```{r}
ggplot(weather_df,aes(x=tmin,y=tmax))+
  geom_point()
```
```{r}
gg_temp_scatterplot=
  weather_df|>
    ggplot(aes(x=tmin,y=tmax,color=name))+
    geom_point(alpha=.25)+ #transparent
    geom_smooth(se=FALSE) #ggplot2 给图加一条平滑趋势线，同时不显示置信区间的阴影
gg_temp_scatterplot
```
show faceting
```{r}
weather_df|>
  ggplot(aes(x=tmin,y=tmax,color=name))+
  geom_point(alpha=.5)+ #transparent
  facet_grid(.~name)#facet_grid(行 ~ 列)
```
```{r}
weather_df|>
  ggplot(aes(x=tmin,y=tmax,color=name))+
  geom_point(alpha=.5)+ #transparent
  facet_grid(cols=vars(name))#facet_grid(行 ~ 列)
```
let's look at something else
```{r}
weather_df|>
  ggplot(aes(x=date,y=tmax,color=name))+#日期和最高温对应
  geom_point(alpha=.5)+
  geom_smooth(se=FALSE)+
  facet_grid(.~name)
```

```{r}
weather_df|>
  filter(name=="CentralPark_NY")|>
  mutate(
    tmax=tmax*9/5 + 32,
    tmin=tmin*9/5 + 32)|>
  ggplot(aes(x=tmax,y=tmin))+
  geom_point()
```
a hex plot
```{r}
weather_df|>
  ggplot(aes(x=tmin,y=tmax))+
  geom_hex()
```
```{r}
weather_df|>
  ggplot(aes(x=tmax,color=name))+
  geom_histogram(position="dodge")#position = "dodge"如果同一个温度区间里有多个 name，把它们的柱子并排放，而不是叠在一起。
```
```{r}
weather_df|>
  ggplot(aes(x=tmax,fill=name))+
  geom_density(alpha=.3)
```
```{r}
weather_df|>
  ggplot(aes(x=tmax,y=tmin, color=name))+
  geom_point()+
  labs(
    title="temperature(Max vs Min)",
    x="Max Temperature(C)",
    y="Min Temperature(C)",
    color="Location",
    caption="Data from NGAA for three weather stations"
  )+
  scale_x_continuous(
    breaks=c(-10,0,15),
    labels=c("-10 C","0","fifteen")#这些刻度显示成什么文字
  )+
  scale_y_continuous(
    trans="sqrt",#平方根转换
    position="right"
  )
```
```{r}
ggp_weather = 
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(size=prcp), alpha = .5) +
  facet_grid(.~name)

ggsave("images/ggp_weather.pdf", ggp_weather, width = 8, height = 5)
```
```{r}
weather_df|>
  ggplot(aes(x=tmax,y=tmin,color=name))+
  geom_point()+
  viridis::scale_color_viridis(#使用 viridis color palette 来给不同的 name 配色
    name="Location",
    discrete=TRUE
  )+
  theme_bw()+
  theme(legend.position="bottom")+
  theme_minimal()#简洁风格
```
```{r}
weather_df|>
  ggplot(aes(x=date,y=tmax,color=name))+
  geom_point()+
  geom_smooth(se=FALSE)+
  labs(
    title="Seasonal trends in Max Temp",
    x="Date",
    y="Max Temp",
    caption="Max daily temp in three weather staions in 2020 and 2021",
    color="Location"
  )+
  viridis::scale_color_viridis(#使用 viridis color palette 来给不同的 name 配色
    name="Location",
    discrete=TRUE
  )+
  theme(legend.position="bottom")+
  theme_minimal()#简洁风格
```
## Two more weird but useful plot things
```{r}
central_park_df=
  weather_df|>
  filter(name=="CentralPark_NY")
molokai_df=
  weather_df|>
  filter(name=="Molokai_HI")
ggplot(molokai_df,aes(x=date,y=tmax,color=name))+
  geom_point()+
  geom_line(data=central_park_df)
```
multiple panels with different plot types
```{r}
ggp_tmax_tmin = 
  weather_df |> 
  ggplot(aes(x = tmax, y = tmin, color = name)) + 
  geom_point(alpha = .5) +
  theme(legend.position = "none")

ggp_prcp_dens = 
  weather_df |> 
  filter(prcp > 0) |> #把没有降水的 observation 排除掉
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .5) + 
  theme(legend.position = "none")

ggp_tmax_date = 
  weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  theme(legend.position = "bottom")

(ggp_tmax_tmin + ggp_prcp_dens) / ggp_tmax_date#先把这两张图组合成第一行
```
## Data Manipulation
start with factors
boxplots
```{r}
weather_df|>
  mutate(name=fct_relevel(name,c("Molokai_HI","CentralPark_NY","Waterhole_WA")))|>
  ggplot(aes(x=name,y=tmax))+
  geom_boxplot()
```

```{r}
weather_df|>
  mutate(name=fct_reorder(name,tmax))|>#从小到大
  ggplot(aes(x=name,y=tmax))+
  geom_boxplot()
```
Make that distribution plot
```{r}
weather_df|>
  select(name,tmax,tmin)|>
  pivot_longer(
    tmax:tmin,
    names_to="observation",
    values_to="temp"
  )|>
  ggplot(aes(x=temp,fill=observation))+#aes:数据变量 → 图上的视觉属性
  geom_density(alpha=.5)+
  facet_grid(.~name)
```
Make an FAS plot
```{r}
pup_data = 
  read_csv("./data/FAS_pups.csv", na = c("NA", ".", ""), skip = 3) |>
  janitor::clean_names() |>
  mutate(
    sex = 
      case_match(
        sex, 
        1 ~ "male", 
        2 ~ "female"))

litter_data = 
  read_csv("./data/FAS_litters.csv", na = c("NA", ".", "")) |>
  janitor::clean_names() |>
  separate(group, into = c("dose", "day_of_tx"), sep = 3)

fas_data = left_join(pup_data, litter_data, by = "litter_number") 

fas_data |> 
  select(sex, dose, day_of_tx, pd_ears:pd_walk) |> 
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome", 
    values_to = "pn_day") |> 
  drop_na() |> 
  mutate(outcome = forcats::fct_reorder(outcome, pn_day, median)) |> 
  ggplot(aes(x = dose, y = pn_day)) + 
  geom_violin() + 
  facet_grid(day_of_tx ~ outcome)
```
```{r}
weather_df |> 
  filter(tmax >= 20, tmax <= 30) |> 
  ggplot(aes(x = tmin, y = tmax, color = name, shape = name)) + 
  geom_point(alpha = .75)
```
```{r}
weather_df =
  weather_df|>
  mutate(
    month = lubridate::floor_date(date, unit = "month")
  )
weather_df|>
  group_by(month)|>
  summarise(#summary function-之后不能再基于它处理了
    n_obs = n(),
    n_days = n_distinct(date))#计算每个月有多少 observations 和多少不同的 days → 最后把这个汇总后的表重新存回 weather_df
```
```{r}
weather_df |>
  group_by(name, month) |>
  summarize(mean_tmax = mean(tmax, na.rm = TRUE)) |> 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax) |> 
  knitr::kable(digits = 1)
```
```{r}
weather_df |>
  group_by(name) |>
  mutate(
    center_tmax = tmax - mean(tmax, na.rm = TRUE)
  ) |>
  ggplot(aes(x = date, y = center_tmax, color = name)) +
  geom_point()
```
```{r}
weather_df |>
  group_by(name, month) |>
  mutate(temp_ranking = min_rank(tmax))#出现数值并列时，使用并列位置中最小的 rank
```
```{r}
weather_df |>
  group_by(name, month) |>
  filter(min_rank(desc(tmax)) < 4)
```
```{r}
weather_df |>
  group_by(name) |>
  mutate(temp_change = tmax - lag(tmax))|>#当前这一行的 tmax − 前一行的 tmax
  summarize(
    temp_change_sd = sd(temp_change, na.rm = TRUE),
    temp_change_max = max(temp_change, na.rm = TRUE))
```
## revisit some examples
import,clean,tidy,etc the pulse data, and compute mean and median BOI score at each visit
```{r}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") |>
  janitor::clean_names() |>
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to="visit",
    names_prefix = "bdi_score_",
    values_to="bdi"
  )|>
  select(id,visit,everything())|>#把 id 和 visit 放到数据集最前面，其他变量全部跟在后面
  mutate(visit=replace(visit,visit=="bl","00m"))#改 visit 变量，把其中等于 "bl" 的值替换成 "00m"
pulse_data
```






















